Assessing exposure to Atlantic Basin tropical storms in United States counties

### 2016-06-14 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Stage IV data info and Lin reference

Currently, it seems that most hurricane track mapping capabilities are restricted to geographical information system software. The introduction of this package expands the mapping potential of hurricane tracks and therefore the assessment of hurricane exposure and impacts to R.
#### Examples of using TS exposure estimates for multiple storms

[@Zandbergen2009] developed an exposure factor based on location (longitude and latitude), county shape, county size, and distance from the coast.

@@

[@Villarini2011] recommends the use of Stage IV precipitation data for the analysis of rainfall distribution in landfalling tropical cyclones. However, the Stage IV dataset only provides data for ...

@@

Some of the other possible sources for estimating rain during tropical storms include ...

@@

Rainfall estimates are likely underestimates for a few reasons. First, they are based on averaging hourly measurements to a daily mean estimate. This averaging would smooth over shorter periods of very extreme rainfall. Further, this data is averaged over multiple grid points within each county and so would not fully reflect very extreme local precipitation (although this might be less of a concern for classifying exposure to a large-scale storm system, like a tropical storm, compared to more fine-scale storms). Finally, this NLDAS data provides a re-analysis that incorporates measured rainfall, using models, etc., to incorporate that observed data into a spatially and temporally continuous dataset of rainfall. However, during extreme storms, the problems with measuring rainfall using [rain monitors] would propogate into the NLDAS data, so although NLDAS would prevent missing values during the storm if monitors are not able to provide data, if monitors are out, rainfall estimates from NLDAS will be based more on models than on observations.

@@

Some storms can be of lower intensity (i.e., on the Saffir-Simpson scale) yet bring dangerous rainfall, including well inland of the storm's landfall. For example, storms including Floyd in 1999, Gaston in 2004, Irene in 2011, and Lee in 2011, have had severe inland impacts, often through extreme rainfall, as post-tropical storms [@Halverson2015]. This rainfall can be particularly severe in the Appalachians [@Halverson2015]. Further, heavier rainfall is more common for storms moving at slower wind speeds [@Chang2013]. Storms with a slower forward speed and larger rainfall area contribute to a longer duration of rainfall, and therefore an increased chance of flood-related damage [@Rezapour2014].

>>>>

Currently, it seems that most hurricane track mapping capabilities are restricted to geographical information system software. The introduction of this package expands the mapping potential of hurricane tracks and therefore the assessment of hurricane exposure and impacts to R.

#### Examples of using TS exposure estimates for multiple storms

[@Zandbergen2009] developed an exposure factor based on location (longitude and latitude), county shape, county size, and distance from the coast. The study assumed symmetrical activity of storm related factors,[@Zandbergen2009] which does not accurately represent storm characteristics once landfall has been made.[@Kruk2010, @Halverson2015]

@@

Given the its high resolution and consistancy with rain gage values, [@Villarini2011] recommends the use of Stage IV precipitation data for the analysis of rainfall distribution in landfalling tropical cyclones. However, the Stage IV dataset only provides data since 2001 [@Lin2005] and is therefore inadequate for use in this analysis.

@@

Some of the other possible sources for estimating rain during tropical storms include ...[Stage IV, TMPA, NEXRAD]

@@

Rainfall estimates are likely underestimates for a few reasons. First, they are based on averaging hourly measurements to a daily mean estimate. This averaging would smooth over shorter periods of very extreme rainfall. Further, this data is averaged over multiple grid points within each county and so would not fully reflect very extreme local precipitation (although this might be less of a concern for classifying exposure to a large-scale storm system, like a tropical storm, compared to more fine-scale storms). Finally, this NLDAS data provides a re-analysis that incorporates measured rainfall, using models, etc., to incorporate that observed data into a spatially and temporally continuous dataset of rainfall. However, during extreme storms, the problems with measuring rainfall using [rain monitors] would propogate into the NLDAS data, so although NLDAS would prevent missing values during the storm if monitors are not able to provide data, if monitors are out, rainfall estimates from NLDAS will be based more on models than on observations. More on NLDAS bias low here [@Villarini2011]. Such bias, though unimportant in our analysis, may play a larger role in the extrapolation to other studies or exposure analysese where the precipitation cutoffs used here may provide a threshold that results in the missclassification of less affected counties as exposed.

@@

Some storms can be of lower intensity (i.e., on the Saffir-Simpson scale) yet bring dangerous rainfall, including well inland of the storm's landfall. For example, storms including Floyd in 1999, Gaston in 2004, Irene in 2011, and Lee in 2011, have had severe inland impacts, often through extreme rainfall, as post-tropical storms [@Halverson2015]. This rainfall can be particularly severe in the Appalachians [@Halverson2015]. Further, heavier rainfall is more common for storms moving at slower forward speeds [@Chang2013]. Storms with a slower forward speed and larger rainfall area contribute to a longer duration of rainfall, and therefore an increased chance of flood-related damage [@Rezapour2014].
